The impact of gas bulk rotation on the Lyman-$\alpha$ line

### 2013-09-12 paper/Analisis.tex
down to section 3.1

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{SpectraDifVelocitiesCentral.png}
  \includegraphics[width=0.45\textwidth]{SpectraDifVelocitiesHOM.png}
\caption{Shape of the \ly line for
    different velocities rotational velocities for spherical
    distributions with $\tau_{H}=10^{7}$. The left (right) panel shows
    the central    (homogeneous) photon distribution. All photons were
    taken into     account regardless of their propagation
    direction.    \label{fig:differentvelocities}}  
\end{figure*}

The central result of this paper is summarized in
Fig.~\ref{fig:differentvelocities} that shows clearly the considerable
impact of rotation on the morphology of the emergent \ly line. Both
panels in the Figure focus on the results for $\tau_{H}=10^{7}$,
showing that the influence of rotation is present both when the
photons are either homogeneously or centrally initialized over the gas
volume.

In the following subsections we characterize the line morphology by
the half-width at half intensity and the peak maxima. In order to
interpret the morphological changes in the line we also report the
median number of scatter for each \ly photon in the
simulation. Finally we measure the bulk escape fraction as a function
of rotational velocity in the presence of dust.

All the results in this section are constructed by taking into
account all outgoing photons regardless of the direction of
propagation. In the next section \ref{subsec:direction} we quantify 
the changes in the observed spectra for observers with different
viewing angles.

@@

The first quantitative conclusion of the effect of rotation in the
\ly line is that double peaks broaden and reduce their intensity
while the line center rises. This produces the impression that, as the
rotational velocity increases, the double peaks are merged into a
single broad emission peak. This is most evident in the highest
rotational velocities for the homogeneously distributed sources
(right panel in Fig.~\ref{fig:differentvelocities}).

@@

Figure \ref{fig:widthvsvelocity} shows how $W_{1/2}$ increases with
rotational velocity. Continuous (dashed) lines connect the results for
homogeneous (central) source distribution. The line width increases
with rotational velocity up to factors of $\sim 2$ at $300$\kms with
respect to the static case. For the temperature $T=10^4$K used in our
radiative transfer calculations the thermal velocity is
$v_{th}=12.8$\kms. For a model with $\tau_{H}$ it means that the
half-width can increase up to $350$\kms (at $V_{\rm max}=300$\kms)
compared to a half-width of $150$\kms in the static case.

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
Figure shows us clearly that in the case of central distributed
sources there is barely any change with rotational velocity in the
range of explored parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $x_{\rm max}=0$.

The transition to a single peak line occurs for systems
where it  becomes easier for the bulk of the photons to escape with the lowest
number of scatterings possible, allowing them not to move very far
from the center of the line. This explains how the single peak stage
is easily achieved in the homogeneous source distribution. In this
case there is a fraction of the photons that are inside a photosphere
region with $\tau_{H,r}\ll \tau_{H}$ where $\tau_{H,r}$ is the optical
depth from the radius of emission to the sphere's surface. This
conditions allows the photons to escape with much less scatterings
compared to the photons emitted at the very center of the sphere. In
turn, it gives the photons less scatterings to be placed far from the
line center. Increasing the rotational velocity $V_{\rm max}$ reduces
the optical depth making the photosphere region effectively larger,
increasing the number of photons escaping close to the lines's
center. For the central case the photosphere is not present, and other
mechanisms must be at play in the steady reduction of the double peak.

For larger values of $\tau_{H}$ the transition to a single peaked
emission occurs for larger values of the rotation velocity. In our
models we find the following correspondence between the optical depth
$\tau_{\rm H}=\{10^5, 10^6, 10^7\}$ and the transitional velocities $V_{\rm
  trans}=\{50-100\kms, 100-200\kms, 200-300\kms\}$ which can only be
constrained to be in the range of velocities in the models that gave a
$x_{\rm max}\neq 0$ and $x_{\rm max}=0$.   In the next section explore the
origin of this trends and show to what extent a low number of
scatterings is correlated with the emergence of a single peak.

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Logarithm of the average number of scatterings as function of
  the velocity. \label{fig:Nscatt}}   
\end{figure}

>>>>

\begin{figure*}
  \includegraphics[width=0.45\textwidth]{SpectraDifVelocitiesCentral.png}
  \includegraphics[width=0.45\textwidth]{SpectraDifVelocitiesHOM.png}
\caption{Shape of the \ly line for
    different velocities rotational velocities for spherical
    distributions with $\tau_{H}=10^{7}$. The left (right) panel shows
    the central (homogeneous) photon distribution. All photons were
    taken into  account regardless of their final direction of propagation.
    \label{fig:differentvelocities}}  
\end{figure*}

The central result of this paper is summarized in Figure
\ref{fig:differentvelocities}. It shows the considerable
impact of rotation on the morphology of the emergent \ly line. Both
panels in the Figure focus on the results for $\tau_{H}=10^{7}$,
showing that the influence of rotation is present both when the
photons are either homogeneously or centrally initialized over the gas
volume.

In the following subsections we characterize the line morphology by
the half-width at half intensity and the peak maxima. In order to
interpret the morphological changes in the line we also report the
median number of scatter for each \ly photon in the
simulation. For the models where dust is included we measure the bulk
escape fraction as a function of rotational velocity. Finally, we make
an estimate of the anisotriopic emission of the models in comparison
with static spheres.

@@

The first quantitative conclusion of the effect of rotation in the
\ly line is that double peaks broaden and reduce their intensity
while the line center rises. This produces the impression that, as the
rotational velocity increases, the double peaks are merged into a
single broad emission peak. This is most evident at the highest
rotational velocities for the homogeneously distributed sources
(right panel in Figure \ref{fig:differentvelocities}).

@@

Figure \ref{fig:widthvsvelocity} shows how $W_{1/2}$ increases with
rotational velocity. Continuous (dashed) lines connect the results for
homogeneous (central) source distribution. For the temperature
$T=10^4$K used in our radiative transfer calculations the thermal
velocity is $v_{th}=12.8$\kms. For a model with $\tau_{H}$ it means
that the half-width can increase up to $350$\kms (at $V_{\rm
  max}=300$\kms) compared to a half-width of $150$\kms in the static
case.

Figure \ref{fig:maximumsvsvelocity} shows the position for the peak
maxima as a function of the rotational velocity $V_{\rm max}$. This
Figure clearly shows that in the case of central distributed
sources there is barely any change with rotational velocity in the
range of explored parameters. However, in the case of
homogeneously emitted sources the maxima position remain close to
constant until beyond some velocity threshold the line becomes single
peaked with $x_{\rm max}=0$ \kms.

The transition to a single peak line seems to occur for systems
where it becomes easier for the bulk of the photons to escape with the lowest
number of scatterings possible, allowing them not to move very far
from the center of the line. This might explain how the single peak stage
is easily achieved in the homogeneous source distribution. In this
case there is a fraction of the photons that are inside a photosphere
region with $\tau_{H,r}\ll \tau_{H}$ where $\tau_{H,r}$ is the optical
depth from the radius of emission to the sphere's surface. This
conditions allows the photons to escape with much less scatterings
compared to the photons emitted at the very center of the sphere. In
turn, it gives the photons less scatterings to be placed far from the
line center.

Increasing the rotational velocity $V_{\rm max}$ reduces
the optical depth making the photosphere region effectively larger,
increasing the number of photons escaping close to the lines's
center. In our models we find the following correspondence between the
optical depth $\tau_{\rm H}=\{10^5, 10^6, 10^7\}$ and the transitional
velocities $V_{\rm   trans}=\{50-100\kms, 100-200\kms, 200-300\kms\}$
which can only be constrained to be in the range of velocities in the
models that gave a $x_{\rm max}\neq 0$ and $x_{\rm max}=0$.

For the central emission the transition to a single peak is never
completed in the range of parameters explored here. The nonappearence
of a single peak phase can be in part explained to the absence of a
photosphere, as is the case in the homogeneous
distribution. Nevertheless, there is a rise in the intensity at the
line center as the rotational velocity increases. This hints that the
encounters with a nonstatic medium are inefficient in changing a
photons' frequency outside the line center.

In the next section we quantify the number of scatterings and explore
to what extent this is related to the emergence of single peak.

@@

\begin{figure}
    \includegraphics[width=0.45\textwidth]{NscattvsVmax.png}
\caption{Logarithm of the average number of scatterings as function of
  the velocity. Continuous/dashed lines represent an
  homogeneous/central distribution of sources. \label{fig:Nscatt}}    
\end{figure}
